\paperTitle

### 2014-05-17 doc/icml2014/undirected.tex
undirected draft finished

\section{Undirected log-linear models}
\label{sec:undirected}

% DONE: need more of a transition from previous sections
Recall that we used tensor factorization and composite likelihood
to estimate the hidden marginals $Z_\sC = \BP(\vh_\sC)$ and $\mOpp{v}{i} = \BP(x_v \mid h_i)$.
For directed models, obtaining the local conditional tables from $Z_\sC$ simply requires
renormalizing $Z_\sC$.
For undirected log-linear models, the canonical parameters cannot be obtained
locally.
Fortunately, we can construct a convex optimization problem
by relating the recovered conditional moments with the expected features of
the log-linear model.

% Define model
As before, let $\sG$ be a graphical model with observed variables $\bx$ and hidden variables $\bh$
(see \figureref{examples-mrf} for an example).
For each clique $\sC \in \sG$, we have a feature vector $\phi_\sC(\vx_\sC, \vh_\sC)$,
where $\vx_\sC$ and $\vh_\sC$ are the observed and hidden variables in $\sC$, respectively.
Let $\phi(\vx, \vh) = \sum_{\sC \in \sG} \theta^\top \phi_\sC(\vx_\sC,\vh_\sC)$ be the global feature vector.
We consider log-linear models of the following form:
\begin{align*}
p_\theta(\vx, \vh) &= \exp\left( \theta^\top \phi(\vx,\vh) - A(\theta) \right),
\end{align*}
where $A(\theta) = \log \sum_{\vx, \vh}  \exp( \theta^\top \phi(\vx,\vh) )$ is the log-partition function.

% FIGURE: example
\begin{figure}
  \centering
  \includegraphics[width=0.6\columnwidth]{figures/mrf.pdf}
%  \subimport{figures/}{mrf.tikz}
  \caption{Example: undirected grid model where each hidden variable has two
  conditionally independent observations.
  We derive a consistent estimate.}
  \label{fig:examples-mrf}
\end{figure}

%As an example, consider an undirected grid model in \figureref{examples-mrf},
%where each hidden variable has two conditionally independent observations.
%In practice, if two features are known to be independent, then they can
  %be separated into independent observed variables.
%Let the parameters of the model be $O \eqdef \Phi(x^a_{i,j}, h_{i,j})
  %= \Phi(x^b_{i,j}, h_{i,j})$ and $T \eqdef \Phi(h_{i,j}, h_{i \pm 1,j \pm 1})$. 
%Note that as this is an undirected model, the clique marginals do not
  %correspond to the parameters of the model.
%To recover the parameters of this model, note that every hidden
%  variable, $h_{ij} \in \sH$ is a bottleneck, with views
%  $x^{a}_{ij},x^{b}_{ij}$ and any other observed variable in the graph. 
%Thus, \LearnMarginals will be able to learn the marginals of each clique
%  and \LearnLogLinear will give us a consistent estimate of the
%  parameters.

Of course, the log-likelihood is non-concave:
\begin{align*}
  L_\text{unsup}(\theta) \eqdef \E_{\vx \sim p^*}[\log \sum_{\vh \in \sH} p_\theta(\vx,\vh)],
\end{align*}
where we are assuming for simplicity that we have infinite data drawn from the true distribution $p^*$.
% PL: this doesn't make sense
%While $\phi_\sC(x,h)$ can be an arbitrary featurization, we assume
%  without loss of generality that $\phi_\sC(x,h)$ are the indicator
%  features $\BI[x,h]$\footnote{Note that $\BI[x,h]$ are sufficient
%  statistics for the model, and hence we do not lose any statistical
%  efficiency in using them.}.

%In this section, we will show that we can extend the results to
  %undirected graphical models parameterized as log-linear latent variable
  %models.
On the other hand, 
if we were able to observe $\vh$, we would optimize the \emph{supervised} likelihood, which is concave:
\begin{align}
L_\text{sup}(\theta) &\eqdef \E_{(\vx,\vh) \sim p^*}[\log p_\theta(\vx,\vh)] \nonumber \\
                     &= \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
\label{eqn:logLinearSupervised}
\end{align}
While we don't have supervised data,
recall that the method of moments provides the hidden marginals $Z_\sC = \BP(\vh_\sC)$
and conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
Assuming our graphical model is in canonical form (\lemmaref{reduction}),
every observed variable is connected to exactly one hidden variable.
Then $Z_\sC$ and $\mOpp{v}{i}$ directly yield marginals for every clique $\BP(\vx_\sC, \vh_\sC)$.
Given these marginals, we can compute the expected local feature vector:
\begin{align}
\label{eqn:logLinearFeatures}
\mu_\sC \eqdef \E[\phi(\vx_\sC,\vh_\sC)] = \sum_{\vx_\sC,\vh_\sC} \BP(\vx_\sC,\vh_\sC) \phi(\vx_\sC,\vh_\sC).
\end{align}
Note that $\{\mu_\sC\}_{\sC \in \sG}$ are exactly the sufficient statistics of the model (\equationref{logLinearSupervised}),
so we can optimize the supervised likelihood objective without actually having any supervised data!
In the finite data regime, the method of moments yields the estimate
$\hat \mu^\text{mom}_\sC$ which approximates the true $\mu_\sC$.
In supervised learning, we obtain a different estimate $\hat\mu^\text{sup}_\sC$ of $\mu_\sC$ based on an empirical average
over data points.
In the limit of infinite data, both estimators converge to $\mu_\sC$.

%We exploit the connection between the moments of the clique potentials and
  %the log-partition function to show that learning parameters is convex
  %given the clique marginals $Z_\sC$. 
%The end result is a consistent estimator for log-linear
  %latent-variable models which satisfy \propertyref{exclusive-views}.
%Finally, we generalize the result to classes when only a subset of the
  %clique marginals can be learned.

%As before, our objective is to find a set of parameters $\hat \theta$
%  that maximize the likelihood of data $\sD$.
%\begin{align}
%  \hat \theta 
%      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \nonumber \\
%      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh) \nonumber \\
%      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} \exp( \theta^\top\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
%\end{align}
%Note that this problem is not concave in $\theta$. 
%
%\note{Consider the supervised setting.
%As we get infinite amounts of data, we get this simple form.
%As MoM is a consistent estimator for marginals, we can optimize this
%  likelihood and it is consistent.
%}

%However, given the moments $\E[ \phi_\sC(\vx,\vh) ] = Z_\sC$, the
%  problem simplifies to,
%\begin{align}
%  \hat \theta 
%  &= \arg\max_{\theta} \sum_{\sC \in \sG} \theta_\sC^\top Z_\sC - A(\theta), \label{eqn:obj}
%\end{align}
%which is concave in $\theta$.
\algorithmref{undirected} summarizes our approach.
In \sectionref{exclusiveViews}, we showed that $\LearnMarginals$ is a consistent
estimator for $Z_\sC$ for any bottlenecked clique in the graphical model. 
Consequently, \algorithmref{undirected}
is also a consistent estimator.\footnote{Of course, the log-linear model must be identifiable.
Identifability is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical.
In general, we can identify the parameters up to the kernel of $\nabla^2 A(\theta)$.}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnParameters (for undirected models)}
  \label{algo:undirected}
  \begin{algorithmic}
    \REQUIRE Conditional moments $\mOpp{v}{i}$ and clique marginals $Z_\sC$.
    \ENSURE Parameters $\theta$
      \STATE Compute expected features $\{\mu_\sC\}_{\sC \in \sG}$ (\equationref{logLinearFeatures}).
      \STATE Optimize \equationref{logLinearSupervised}.
    \end{algorithmic}
\end{algorithm}

\subsection{Pseudolikelihood}
\label{sec:pseudolikelihood}

Unlike for directed models, in the undirected case,
graphs with high treewidth present a challenge
for even fully-supervised maximum likelihood due to the intractability of the log-partition function $A(\theta)$.
%One can employ various variational approximations of $A(\theta)$ \cite{wainwright08varinf},
This motivates looking at pseudolikelihood objective of \citep{besag75pseudo} for a computationally more
efficient solution.
The pseudolikelihood objective is a sum over the log-probability of each variable $a$ given its neighbors $\sN(a)$:
\begin{align}
  \label{eqn:pseudolikelihood}
L_\text{pseudo}(\theta) \eqdef \E_{(\bx,\bh) \sim p^*}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid \sN(a))\right].
\end{align}
In the fully-supervised setting, it is well-known that pseudolikelihood provides computationally efficient consistent estimates,
albeit at the cost of statistical efficiency.
Coincidentally, this is the same high-level motivation for using method of moments in the first place.

Let $\phi_{a,\sN(a)}(a, \sN(a)) = \sum_{\sC \ni a} \phi_\sC(\bx_\sC,\bh_\sC)$
denote the sum over cliques $\sC$ that contain $a$; note that $\phi_{a, \sN(a)}$ only depends on $a$ and its neighbors $\sN(a)$.
We can write each conditional log-likelihood as:
\begin{align*}
p_\theta(a \mid \sN(a)) = \exp(\theta^\top\phi_{a,\sN(a)}(a, \sN(a)) - A_a(\theta; \sN(a))),
\end{align*}
where the conditional log-partition function $A_a(\theta; \sN(a)) = \log \sum_{a'} \exp(\theta^\top\phi_{a,\sN(a)}(a', \sN(a)))$
involves marginalizing only over a single variable.

If we knew $\mu_{a,\sN(a)} \eqdef \E[\phi_{a,\sN(a)}(a,\sN(a))]$,
then we would be able to optimize the pseudolikelihood objective again without
having access to any labeled data.
%The pseudolikelihood also replaces the log-partition function
%  $A(\theta)$ with a sum over the variables in the neighborhood:
%  $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)}
%  \exp(\theta^\top\phi_{\sN(a)}(a'))$.
Note that $\{a\}\cup\sN(a)$ in general does not form a clique.
Fortunately, \lemmaref{bottleneck-views} holds for all biconnected subgraphs.
We can break $\{a\}\cup\sN(a)$ into a collection of biconnected subgraphs 
  % PL: this is not necessary
  %thus the uniformly bottlenecked assumption alone is not sufficient to
  %guarantee exclusive views for each variable in that subgraph.
%However, when we do have exclusive views for each neighborhood, as in
 (e.g. \figureref{examples-mrf}).
  Therefore, we can adapt our algorithm ($\LearnMarginals$) to estimate all marginals of the form $Z_{a,\sN(a)} \eqdef \Pr(a,\sN(a))$,
  from which we can compute the features $\mu_{a,\sN(a)}$,
  and optimize \equationref{pseudolikelihood} to obtain a consistent estimate of $\theta$.

% Arun: WRONG
% The pseudolikelihood also replaces the log-partition function $A(\theta)$
% with a sum over the variables in the neighborhood: $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)} \exp(\theta^\top\phi_{\sN(a)}(a'))$,
% which cannot be computed from just the hidden clique marginals $Z_\sC$,
% since the neighbors $\sN(a)$ in general do not form a clique.
% We adapt our earlier algorithm ($\LearnClique$) to estimate the
%   marginals over all neighborhoods $Z_{\sN(a)}$ by noting that
%   \lemmaref{bottleneck-views} guarantees exclusive views for any connected
%   subgraph of $\sG$, not just cliques.

While this method is only viable for graphs of low degree, i.e.\@ when $|\sN(a)|$ is small,
it can handle graphs with high treewidth.
For example, in the grid model, the degree is only 4.
Note that neither sample nor computational complexity depends on the treewidth.

\paragraph{Learning with partial moment constraints.}

In general, it might be case that we only have exclusive views for a subset of the cliques.
In this case, we can estimate the expected features $\mu_\sC$ for those cliques
  and use posterior regularization \citep{graca08em} or the generalized expectation criteria \citep{mann08ge}
  to encourage $\E_{p_\theta}[\phi(\vx_\sC,\vh_\sC)]$ to match $\mu_\sC$.
The resulting objective functions are still non-convex, but we expect
  local optima to be alleviated.

%but in practice we observed that the extra information $\mu_\sC$ helps alleviate local optima.
%We can then integrate these constraints using posterior regularization \citep{graca08em}:
%\begin{align}
%L_\text{pr}(\theta) \eqdef L_\text{unsup}(\theta) + C\min_{q \in \sQ} \text{KL}(q(\vx,\vh) \| p_\theta(\vx,\vh)).
%\end{align}
%The main idea is to regularize the parameters towards values that produce the 
%While the optimization problem \eqref{eqn:obj-ml} is non-convex, we
  %expect that the measurements aid recover of parameters, and indeed see
  %this is true for experiments on simulated data.
%\todo{Include figure}

%Posterior regularization optimizes the following objective,
%\begin{align*}
%  \hat \theta &= \arg\min_{\theta} \arg\min_{q \in \sQ} \KL( q(\vx,\vh) \| p_{\theta}(\vx, \vh) ),
%\end{align*}
%where $q$ is a marginal distribution that belongs in constrained set
%$\sQ$. In our case, $\sQ$ is constrained to agree with the partial
%moments we recovered using \LearnMarginals, i.e. 
%\begin{align*}
%  \sQ &= \{ q(\vx,\vh) \given \forall \sC \in \sP ~ \E_{q(\vx,\vh)}[\phi_\sC(\vx, \vh)] = Z_\sC \}.
%\end{align*}

%In practice, we would like to regularize for error in our
%  estimates of $Z_\sC$; to do so, we use the measurements model of
%  \citet{liang09measurements} which defines a Bayesian model for the noise
%  in the marginals $Z_\sC$. The resultant optimization problem is,
%\begin{align*}
%  \hat\theta 
%  &= \arg\min_\theta - \sum_{x \in \sD} \log p_\theta(x) + \half \eta_\theta \| \theta \|^2 \\
%  &\quad + \half \eta_\beta \| Z_\sC - \sum_{\vx \in \sD} \E_{\theta}[\phi_\sC(\vx,\vh)] \|^2,
%\end{align*}
%where $\eta_\theta$ and $\eta_\beta$ are regularization parameters.

%
%However, the objective \equationref{obj} requires us to compute the
%  partition function, $A(\theta)$ and its gradient. 
%This scales with the treewidth of the model, which is the number
%  of rows/columns of the grid in this example, and hence can be
%  intractable.
%A more computationally efficient, but still consistent, alternative is
%to optimize the pseudo-likelihood \todo{how do I write this out}.

>>>>

\section{Recovering parameters}
\label{sec:undirected}

% Transition
We have thus far shown how to recover the conditional moments
$\mOpp{v}{i} = \BP(x_v \mid h_i)$ for each exclusive view $x_v$ of each hidden variable $h_i$,
as well as the hidden marginals $Z_S = \BP(\vh_S)$ for each bidependent subset of hidden variables $S$.
Now all that remains to be done is to recover the parameters.

Since our graphical model is in canonical form (\assumptionref{canonical}),
all cliques $\sC \in \sG$
either consist of hidden variables $\vh_\sC$ or are of the form $\{x_v,h_i\}$.
The key observation is that the clique marginals are actually sufficient
statistics of the model $p_\theta$.
How we turn these clique marginals $\{ \Pr(\bx_\sC,\bh_\sC) \}_{\sC \in \sG}$
into parameters $\theta$ depends on the exact model parametrization.

For directed models, the parameters are simply the local conditional tables
$p_\theta(a \mid \Pa(a))$ for each clique $\sC = \{a\} \cup \Pa(a)$.
These conditional distributions can be obtained by simply normalizing $Z_\sC$
for each assignment of $\Pa(a)$.

%by relating the recovered conditional moments with the expected features of
%the log-linear model.

% Define model
%As before, let $\sG$ be a graphical model with observed variables $\bx$ and hidden variables $\bh$
%(see \figureref{examples-mrf} for an example).
%For each clique $\sC \in \sG$, we have a feature vector $\phi_\sC(\vx_\sC, \vh_\sC)$,
%where $\vx_\sC$ and $\vh_\sC$ are the observed and hidden variables in $\sC$, respectively.
%Let $\phi(\vx, \vh) = \sum_{\sC \in \sG} \theta^\top \phi_\sC(\vx_\sC,\vh_\sC)$ be the global feature vector.
%We consider log-linear models of the following form:
%\begin{align*}
%p_\theta(\vx, \vh) &= \exp\left( \theta^\top \phi(\vx,\vh) - A(\theta) \right),
%\end{align*}
%where $A(\theta) = \log \sum_{\vx, \vh}  \exp( \theta^\top \phi(\vx,\vh) )$ is the log-partition function.

% FIGURE: example
\begin{figure}
  \centering
  \includegraphics[width=0.6\columnwidth]{figures/mrf.pdf}
%  \subimport{figures/}{mrf.tikz}
  \caption{Example: undirected grid model where each hidden variable has two
  conditionally independent observations.
  This model has high treewidth,
  but we can estimate it efficiently using pseudolikelihood.}
  \label{fig:examples-mrf}
\end{figure}

%Of course, the log-likelihood is non-concave:
%\begin{align*}
  %L_\text{unsup}(\theta) \eqdef \E_{\vx \sim p_{\theta^*}}[\log \sum_{\vh \in \sH} p_\theta(\vx,\vh)],
%\end{align*}
%where we are assuming for simplicity that we have infinite data drawn from the true distribution $p_{\theta^*}$.
For undirected log-linear models, the canonical parameters $\theta$ cannot be obtained locally,
but we can construct a global convex optimization problem to solve for $\theta$.
Suppose we were able to observe $\vh$.
Then we could optimize the \emph{supervised} likelihood, which is concave:
\begin{align}
  L_\text{sup}(\theta)
  & \eqdef \E_{(\vx,\vh) \sim p_{\theta^*}}[\log p_\theta(\vx,\vh)] \nonumber \\
  & = \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
\label{eqn:logLinearSupervised}
\end{align}
Of course we don't have supervised data,
but we do have the marginals $\Pr(\vx_\sC,\vh_\sC)$,
from which we can easily compute the expected features:
%recall that the method of moments provides the hidden marginals $Z_\sC = \BP(\vh_\sC)$
%and conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
%Assuming our graphical model is in canonical form (\lemmaref{reduction}),
%every observed variable is connected to exactly one hidden variable.
%Then $Z_\sC$ and $\mOpp{v}{i}$ directly yield marginals for every clique $\BP(\vx_\sC, \vh_\sC)$.
%Given these marginals, we can compute the expected local feature vector:
\begin{align}
\label{eqn:logLinearFeatures}
\mu_\sC \eqdef \E[\phi(\vx_\sC,\vh_\sC)] = \sum_{\vx_\sC,\vh_\sC} \BP(\vx_\sC,\vh_\sC) \phi(\vx_\sC,\vh_\sC).
\end{align}
%Note that $\{\mu_\sC\}_{\sC \in \sG}$ are exactly the sufficient statistics of the model (\equationref{logLinearSupervised}),
Therefore, we can optimize the supervised likelihood objective without actually
having any supervised data!
In the finite data regime, the method of moments yields the estimate
$\hat \mu^\text{mom}_\sC$ which approximates the true $\mu_\sC$.
In supervised learning, we obtain a different estimate $\hat\mu^\text{sup}_\sC$ of $\mu_\sC$ based on an empirical average
over data points.
In the limit of infinite data, both estimators converge to $\mu_\sC$.

%\algorithmref{undirected} summarizes our approach.
%In \sectionref{exclusiveViews}, we showed that $\LearnMarginals$ is a consistent
%estimator for $Z_\sC$ for any bottlenecked clique in the graphical model. 
%Consequently, \algorithmref{undirected}
%is also a consistent estimator.\footnote{Of course, the log-linear model must be identifiable.
%Identifability is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical.
%In general, we can identify the parameters up to the kernel of $\nabla^2 A(\theta)$.}

%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnParameters}
  \label{algo:undirected}
  \begin{algorithmic}
    \REQUIRE Conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$
             and hidden marginals $Z_S = \Pr(\bh_S)$.
    \ENSURE Parameters $\theta$.
    \IF{$\sG$ is directed} \STATE Normalize each $\Pr(a,\Pa(a))$. \ENDIF
    \IF{$\sG$ is undirected with low treewidth}
      \STATE Compute expected features $\{\mu_\sC\}_{\sC \in \sG}$ (\equationref{logLinearFeatures}).
      \STATE Optimize full likelihood (\equationref{logLinearSupervised}).
    \ENDIF
    \IF{$\sG$ is undirected with high treewidth}
      \STATE Compute expected features $\{\mu_S\}_{S \in \sB}$ (\equationref{logLinearFeatures}).
      \STATE Optimize pseudolikelihood (\equationref{pseudolikelihood}).
    \ENDIF
    \end{algorithmic}
\end{algorithm}

% Learning with partial moment constraints
\paragraph{Remark} If we have exclusive views for only a subset of the cliques,
we can still obtain the expected features $\mu_\sC$ for those cliques
  and use posterior regularization \citep{graca08em}, measurements \citep{liang09measurements}, or generalized
  expectation criteria \citep{mann08ge}
  to encourage $\E_{p_\theta}[\phi(\vx_\sC,\vh_\sC)]$ to match $\mu_\sC$.
The resulting objective functions would be non-convex, but we expect
  local optima to be less of an issue.
  %How to characterize this precisely is an interesting direction for future work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pseudolikelihood}
\label{sec:pseudolikelihood}

While we now have a complete algorithm for estimating directed and undirected
models, optimizing the full likelihood (\equationref{logLinearSupervised})
can still be computationally intractable for undirected models with high treewidth
due to the intractability of the log-partition function $A(\theta)$.
One can employ various variational approximations of $A(\theta)$ \cite{wainwright08varinf},
but these generally lead to inconsistent estimates of $\theta$.
We thus turn to an older idea of pseudolikelihood \citep{besag75pseudo}.
The pseudolikelihood objective is a sum over the log-probability of each
variable $a$ given its neighbors $\sN(a)$:
\begin{align}
  \label{eqn:pseudolikelihood}
  L_\text{pseudo}(\theta) \eqdef \E_{(\bx,\bh) \sim p_{\theta^*}}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid \sN(a))\right].
\end{align}
In the fully-supervised setting, it is well-known that pseudolikelihood
provides consistent estimates which are computationally efficient
but less statistically efficiency.
(Coincidentally, this is the same high-level motivation for using method of
moments in the first place.)

Let $\phi_{a,\sN(a)}(a, \sN(a)) = \sum_{\sC \ni a} \phi_\sC(\bx_\sC,\bh_\sC)$
denote the sum over cliques $\sC$ that contain $a$; note that $\phi_{a, \sN(a)}$ only depends on $a$ and its neighbors $\sN(a)$.
We can write each conditional log-likelihood from \equationref{pseudolikelihood} as:
\begin{align*}
p_\theta(a \mid \sN(a)) = \exp(\theta^\top\phi_{a,\sN(a)}(a, \sN(a)) - A_a(\theta; \sN(a))),
\end{align*}
where the conditional log-partition function $A_a(\theta; \sN(a)) = \log \sum_{a'} \exp(\theta^\top\phi_{a,\sN(a)}(a', \sN(a)))$
involves marginalizing only over the single variable $a$.

If we knew $\mu_{a,\sN(a)} \eqdef \E[\phi_{a,\sN(a)}(a,\sN(a))]$,
then we would be able to optimize the pseudolikelihood objective again without
having access to any labeled data.
However, $\{a\}\cup\sN(a)$ does not always have exclusive views
(for example, consider $a = h_1$ and $\sN(a) = \{h_2,h_3,h_4\}$ in \figureref{examples-tree}).

However, we can decompose $\{a\}\cup\sN(a)$ as follows:
conditioning on $a$ partitions $\sN(a)$ into independent subsets;
let $\sB(a)$ be the collection of these subsets,
which we will call \emph{sub-neighborhoods}.
For example, $\sB(h_1) = \{ \{h_2\},\{h_3\},\{h_4\} \}$ in \figureref{examples-tree}
and $\sB(h_{2,2}) = \{ \{ h_{1,2}, h_{2,3}, h_{3,2}, h_{2,1} \} \}$ contains a single sub-neighborhood.

A key observation is that for each $B \in \sB(a)$, each $\{a\} \cup B$ is bidependent:
conditioning on $a$ does not introduce new independencies within $B$ by construction of $\sB(a)$,
and conditioning on any $b \in B$ does not either since every other $b' \in B \backslash \{b\}$
is connected to $a$.
Assuming $\sG$ is bottlenecked, by \lemmaref{bottleneck-views},
we have that $\{a\} \cup B$ has exclusive views.
Hence, we can recover $\Pr(a,B)$ for each $a$ and $B \in \sB(a)$.
Based on conditional independence of the sub-neighborhoods $B$ given $a$,
we have that $\Pr(a, \sN(a)) = \Pr(a) \prod_{B \in \sB(a)} \Pr(B \mid a)$.
This allows us to estimate the expected features $\mu_{a,\sN(a)}$ and use these quantities
in the optimization of the pseudolikelihood objective.

%  Therefore, we can adapt our algorithm ($\LearnMarginals$) to estimate all
%  marginals of the form $Z_{a,\sN(a)} \eqdef \Pr(a,\sN(a))$,
%  from which we can compute the features $\mu_{a,\sN(a)}$,
%  and optimize \equationref{pseudolikelihood} to obtain a consistent estimate of $\theta$.

Note that our pseudolikelihood-based approach
does depend exponentially on the size of the sub-neighborhoods,
which could be larger than the largest clique size.
Therefore,
for this approach to be effective,
each node essentially should have low degree or locally exhibits a lot of
conditional independence.
%While this method is only viable for graphs with small sub-neighboof low
%degree, i.e.\@ when $|\sN(a)|$ is small,
On the positive side, we can handle graphical models with high treewidth;
neither sample nor computational complexity depends on the treewidth.
For example, in the grid model, the treewidth is linear in the size of the grid,
but it is the degree (only 4 in this case) that governs the computational complexity.
